### 2023-09-29 sections/conclusion.tex

\section{Conclusion and Further Work}

\begin{itemize}
\item implementation for ZX
\item computational linguistics
\item traced categories, hypergraphs for traced comonoids and full e-graphs
\item closed categories, hierarchical hypergraphs and lambda-calculus (example)
\end{itemize}