### 2023-12-19 sections/soundness_and_completeness.tex
TODOs added

% !TEX root = ../main.tex

\section{Soundness and Completeness}

\update{Below I kind of use $\llbracket - \rrbracket$ to define itself when using $\llbracket \mathcal{E} \rrbracket$.}
In this section, we will build a correspondence between terms in a semilattice enriched $\textsf{PROP}^{+}(\Sigma)$ and morphisms in $\WellTypedMdaEcospans$.
Similarly to $\textsf{PROP}^{+}(\Sigma)$ where terms are defined up to semilattice enrichment and SMC equations we will say that two morphisms $G$ and $F$ in $\WellTypedMdaEcospans$ are equal when $G \Rrightarrow{}_{\mathcal{S}} F$. 
As SMC equations are absorbed by e-hypergraph representation we only quotient by $\mathcal{S}$. 
We are now ready to build a semilattice-enriched SMC functor

\[
[\cdot] : \textsf{PROP}^{+}(\Sigma) \to \sfrac{\WellTypedMdaEcospans}{\mathcal{S}}
\]
where $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is $\WellTypedMdaEcospans$ quotiented by structural rewrite rules which makes this category into a semilattice-enrhiched one.
By freeness, it suffices to define how the functor acts on generators \update{Identities and generators from the signature?}.
% As $\textsf{PROP}(\Sigma)$ and $\textsf{PROP}^{+}(\Sigma)$ essentially have the same set of generators we can reuse the functor $\llbracket \cdot \rrbracket: \textsf{PROP}(\Sigma) \to \MdaCospans$ from~\cite{Frobenius2} and then follow the inclusion $\mathcal{I}$ defined in~\ref{functor:I}.
Everything else follows by defining

\begin{enumerate}
    \item $[ f;g ] = [ f ] ; [ g ]$  for $f,g \in \Sigma$;
    \item $[ f \otimes g ] = [ f ] \otimes [ g ]$  for $f,g \in \Sigma$;
    \item $[ f + g ] = [ f ] + [ g ]$  for $f,g \in \Sigma$;
    \item \update{Symmetry?}
\end{enumerate}

this also makes the functor into a symmetric monoidal semilattice-enriched SMC one on the nose.
% \update{This essentially means that $[\cdot] = \llbracket \cdot \rrbracket ; \mathcal{I}$ plus three equations above.}

\update{Anything about soundness? I.e. if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$? Because $f = g$ is modulo semilattice-enriched SMC equations we have to show that $\llbracket f \rrbracket \Rrightarrow \llbracket g \rrbracket$. Doesn't it require the soundness of rewriting?
}

\begin{theorem}
    If $f = g$ in $PROP^{+}(\Sigma)$ then $[f] \Rrightarrow_{\mathcal{S}} [g]$ in $\WellTypedMdaEcospans$.
    \begin{proof}
        \update{
            By induction on the number of applied equations to turn $f$ into $g$?
            E.g. $f = g$ in one step essentially means that $f = l$ for some $l$ in~\ref{fig:string-equations} and $g = r$ (or the other way around).
            Then we need to show that $[f] \Rrightarrow_{\langle l, r \rangle \in \mathcal{S}} [g']$ and $[g'] = [g]$.
            }
        \question{Can we use normal form here as well? $f = g$ means there exists $h$ such that $f =_{SMC^{+}} h$ and $h =_{SMC^{+}}$ and $h$ is of the form $\phi_1 + \ldots + \phi_n$?
                  Then we need to show that $[f] \Rrightarrow_{\mathcal{S}} [h]$ and $[h] \Rrightarrow_{\mathcal{S}} [g]$.
        }
    \end{proof}
\end{theorem}

\begin{corollary}
    $[\cdot] : PROP^{+}(\Sigma) \to \sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is sound in a sense that
    if $f = g$ in $PROP^{+}(\Sigma)$ then $[f] = [g]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$
\end{corollary}

\begin{proposition}
    Functor $\llbracket \cdot \rrbracket; \mathcal{I}$ is faithful.
    \begin{proof}
        Consider the commutative diagram below \update{Is it obvious that the diagram commutes?}.

        % https://q.uiver.app/#q=WzAsNCxbMCwwLCJcXHRleHR7UFJPUH0oXFxTaWdtYSkiXSxbMiwwLCJcXHRleHR7UFJPUH1eeyt9KFxcU2lnbWEpIl0sWzAsMiwiXFxNZGFDb3NwYW5zIl0sWzIsMiwiXFxzZnJhY3tcXFdlbGxUeXBlZE1kYUVjb3NwYW5zfXtcXG1hdGhjYWx7U319Il0sWzAsMSwiXFxtYXRoY2Fse0p9Il0sWzAsMiwiXFxsbGJyYWNrZXQgXFxjZG90IFxccnJicmFja2V0IiwyXSxbMSwzLCJbXFxjZG90XSJdLFsyLDMsIlxcbWF0aGNhbHtJfSIsMl1d
        \[
        \begin{tikzcd}
            {\text{PROP}(\Sigma)} && {\text{PROP}^{+}(\Sigma)} \\
            \\
            \MdaCospans && {\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}}
            \arrow["{\mathcal{J}}", from=1-1, to=1-3]
            \arrow["{\llbracket \cdot \rrbracket}"', from=1-1, to=3-1]
            \arrow["{[\cdot]}", from=1-3, to=3-3]
            \arrow["{\mathcal{I}}"', from=3-1, to=3-3]
        \end{tikzcd}
        \]
    \end{proof}

    $\llbracket \cdot \rrbracket$ is faithfull by Corollary 4.3 in~\cite{Frobenius}.
    $\mathcal{J}$ is an inclusion that maps terms in $\textsf{PROP}(\Sigma)$ to SMC fragment of $\textsf{PROP}^{+}(\Sigma)$ and is automatically faithful. 
    Similarly, $\mathcal{I}$ is faithful because it is inclusion.
    Then, by composition, $\llbracket \cdot \rrbracket; \mathcal{I}$ is faithful. 
\end{proposition}

\begin{theorem}[$\llbracket - \rrbracket$ is faithful]
\begin{proof}
Essentially we need to show that if $\llbracket f \rrbracket = \llbracket g \rrbracket$ then $f = g$ which is equivalent to if $f \not = g$ then $\llbracket f \rrbracket \not = \llbracket g \rrbracket$.
First, we note that we have a \question{weak (modulo commutativity)} normal form for any term \update{string diagram?} in $PROP^{+}(\Sigma)$ which is achieved by applying distributivity for tensor and composition as well as unpacking equations, that is, every term is representable as in Figure~\ref{fig:normal_form_string} where $\phi_1,\;\ldots\;,\phi_n$ are pure SMC terms.
\update{It is not actaully a `normal' form because we can still apply SMC equations, so it should be normal modulo SMC and commutativity}.

\begin{figure}
    \[
    \scalebox{0.75}{
        \tikzfig{soundness_and_completeness/normal_form_string}
    }    
    \]
    \caption{Weak normal form of a term in $PROP^{+}(\Sigma)$}
    \label{fig:normal_form_string}
\end{figure}

Then we get a corresponding cospan for such weak normal form in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ as in Figure~\ref{fig:normal_form_ehyp}

\begin{figure}

\[
\scalebox{0.75}{
\tikzfig{soundness_and_completeness/normal_form}
}
\]
\caption{Interpretation of the weak normal form in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$}
\label{fig:normal_form_ehyp}
\end{figure}

where all $A_1\; \ldots \;A_n$ do not contain any hierarchical edges.
We will write such forms as $\phi_1 + \ldots \phi_n$ and $A_1 + \ldots A_n$ respectively.
Now, assume that $f = \phi_1 + \ldots + \phi_n$ and $g = \psi_1 + \ldots + \psi_m$.
Then, by functoriality, $[f] = [\phi_1] + \ldots + [\phi_n] = A_1 + \ldots + A_n$ and $g = [\psi_1] + \ldots + [\psi_m] = B_{1} + \ldots + B_{m}$. 
$f \not = g$ gives us two options.

\begin{itemize}
    \item $n \not = m$. Then $[f] \not = [g]$ as they are simply not isomorphic cospans of e-hypegraphs as they have different number of summands.
    \item $n = m$. Then there exists a permutation $\sigma$ and index $i$ such that $\phi_i \not = \psi_{\sigma(i)}$. 
    Remember that these are just SMC terms and by faithfulness of $\llbracket \cdot \rrbracket;\mathcal{I}$ we get $A_i \not = B_{\sigma(i)}$ and $[f] \not = [g]$.
\end{itemize}

\end{proof}
\end{theorem}

\question{Is proof ok?}

% \begin{theorem}

%     Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$ and $\langle l, r \rangle \in \mathcal{R}$.
%     $a \Rightarrow{}^{+}_{\langle l, r \rangle} b$ iff $[a] \Rrightarrow{}^{+}_{\langle [l], [r] \rangle} [b]$.
%     \begin{proof}
%         $a \Rrightarrow^{+}_{\langle l, r \rangle} b$ means that either
%         \begin{itemize}
%             \item $a$ and $b$ are decomposable as
%                   \[
%                     \scalebox{0.75}{
%                         \tikzfig{combinatorial_semantics/prop_rewriting}
%                     }
%                   \]
%                   \question{Here we can probably just reuse the proof going to $\ulcorner a \urcorner$ and then the result follows by functoriality of $[\cdot]$ (or $\llangle \cdot \rrangle$)}
%             \item or
%                  \[
%                     \scalebox{0.75}{
%                         \tikzfig{combinatorial_semantics/prop_plus_rewriting}
%                     }   
%                  \]
%                  and $a' \Rightarrow_{\langle l,r \rangle} b'$. This gives us the following decomposition in cospans
%                  % https://q.uiver.app/#q=WzAsMjcsWzAsNCwibSJdLFsxLDQsIkFfMSJdLFsyLDQsImkgKyBrIl0sWzMsNCwiOyJdLFs0LDAsImkiXSxbNSwwLCJYXzEiXSxbNiwwLCJqIl0sWzUsMiwiXFx2ZG90cyJdLFs1LDQsIkEnIl0sWzQsNCwiaSJdLFs2LDQsImoiXSxbNSwxLCIrIl0sWzUsMywiKyJdLFs1LDUsIisiXSxbNSw2LCJcXHZkb3RzIl0sWzUsNywiKyJdLFs0LDgsImkiXSxbNSw4LCJYX3ciXSxbNiw4LCJqIl0sWzcsNCwiXFxvdGltZXMiXSxbOCw0LCJrIl0sWzksNCwiayJdLFsxMCw0LCJrIl0sWzExLDQsIjsiXSxbMTIsNCwiaiArIGsiXSxbMTMsNCwiQV8yIl0sWzE0LDQsIm4iXSxbMCwxXSxbMiwxXSxbNCw1XSxbNiw1XSxbOSw4XSxbMTAsOF0sWzE4LDE3XSxbMTYsMTddLFsyMCwyMV0sWzIyLDIxXSxbMjYsMjVdLFsyNCwyNV1d
%                 \[
%                     \adjustbox{scale=0.5, center}{
%                 % \scalebox{0.75}{
%                     \begin{tikzcd}
%                     &&&& i & {X_1} & j \\
%                     &&&&& {+} \\
%                     &&&&& \vdots \\
%                     &&&&& {+} \\
%                     m & {A_1} & {i + k} & {;} & i & {A'} & j & \otimes & k & k & k & {;} & {j + k} & {A_2} & n \\
%                     &&&&& {+} \\
%                     &&&&& \vdots \\
%                     &&&&& {+} \\
%                     &&&& i & {X_w} & j
%                     \arrow[from=5-1, to=5-2]
%                     \arrow[from=5-3, to=5-2]
%                     \arrow[from=1-5, to=1-6]
%                     \arrow[from=1-7, to=1-6]
%                     \arrow[from=5-5, to=5-6]
%                     \arrow[from=5-7, to=5-6]
%                     \arrow[from=9-7, to=9-6]
%                     \arrow[from=9-5, to=9-6]
%                     \arrow[from=5-9, to=5-10]
%                     \arrow[from=5-11, to=5-10]
%                     \arrow[from=5-15, to=5-14]
%                     \arrow[from=5-13, to=5-14]
%                 \end{tikzcd}}
%                 \]
%                 and there is a convex match of $L$ in $A'$.
%                 It is then also down-closed and conflict reflecting as $A'$ is not hierarchical.
%                 Need to show that arising pushout complement is a boundary complement.
%                 \question{A problem is that the above terms does not decompose into a form of $A';C$ where $C$ is a pushout complement.
%                 This happens because we take composition along the external interfaces and this $A'$ is within a box, so it is not a part of an external interface.
%                 }
%                 \Aleksei{discuss this. We can simplify PROP+ rewriting to just a box, but still have a problem with context C}
%         \end{itemize}
%         \update{It is tempting to reuse a similar result for $\llbracket \cdot \rrbracket$ because according to our definition $f \Rrightarrow_{\mathcal{R}^{+}}$ means that there is an SMC piece $f'$ inside some box that rewrites to $g'$.
%             But we only have this result for $\llangle \cdot \rrangle$ for terms with bent wires which is a functor to $PROP + Frob$.
%         }
%     \end{proof}
% \end{theorem}

We will next consider functors
\[
    \textsf{PROP}(\Sigma, \mathcal{E}) \xrightarrow{\mathcal{J}} \textsf{PROP}^{+}(\Sigma, \mathcal{E}) \xrightarrow{\llangle - \rrangle} \sfrac{\WellTypedMdaEcospans}{\mathcal{S},\mathcal{E}}
\]
where $\mathcal{E}$ is a set of equations of the corresponding SMT.
\begin{theorem}[$\llangle - \rrangle$ is a functor]
\end{theorem}

\begin{theorem}[$\llangle - \rrangle$ is faithful]
\end{theorem}

\update{Or we can define rewriting for string diagrams in $\textsf{PROP}^{+}(\Sigma)$ based on rewriting in $\textsf{PROP}(\Sigma)$ and then say that $f \xrightarrow{}_{\mathcal{R}} g$ iff $[f] [g]$}

\begin{definition}[Rewriting in $\textsf{PROP}^{+}(\Sigma)$]

Recall the definition of rewriting in some $\textsf{PROP}(\Sigma) = \mathbb{A}$.
We say that $a \in \mathbb{A}$ rewrites into $b \in \mathbb{A}$ under a rewrite rule $\langle l, r \rangle$ where $l,r \in \mathbb{A}$ if $a$ and $b$ are decomposable as below.
\[
\scalebox{0.75}{
\tikzfig{combinatorial_semantics/prop_rewriting}
}
\]
and we will use notation $a \Rightarrow_{\langle l, r \rangle} b$.

Now consider an enriched prop $\mathbb{A}^{+}$ over the signature $\Sigma$ such that there is a functor $\mathcal{J} : \mathbb{A} \to \mathbb{A}^{+}$.
We than say that $a \in \mathbb{A}^{+}$ rewrites into $b \in \mathbb{A}^{+}$ under a rewrite rule $\langle l, r \rangle$ where $l,r \in \mathcal{J}(\mathbb{A})$ if $a$ and $b$ are decomposable as below.

\[
\scalebox{0.75}{
\tikzfig{combinatorial_semantics/prop_plus_rewriting}
}
\]

and $a' \Rightarrow_{\langle l, r \rangle} b'$.
We will use notation $a \Rightarrow_{\langle l, r \rangle}^{+} b$.
\end{definition}

\begin{theorem}
    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$.
\end{theorem}