### 2024-03-19 Reviews-response.tex
reviews-response draft

%!LW recipe=pdflatex
\documentclass{article}

\usepackage{tikz-cd}
\usepackage{xargs}
\usepackage{amsmath}

\input{preamble}
\input{macros}
\input{sample.tikzstyles}
\input{hypergraph.tikzstyles}
\input{hypergraph.tikzdefs}

\newcommand\id{\textsf{id}}
\newcommand\sym{\textsf{sym}}

\begin{document}

\title{LICS24 Reviews Response}
\author{}

\maketitle

\section*{Review 1}
\paragraph{About theorem 6.5 (the central result of the paper: full completeness)}
\textit{I am doubtful about the statement because it seems to me that $[sym] = [id]$,
hence the functor $[-]$ is not faithful.
Here is a sketch of the argument: $[\sym]$ is a e-hypergraph with two vertices
where the inputs and the outputs are inverted. Now we embed it into a
hierarchical edge by applying the bottom left equality of Figure 3.
We get a cospan with 2 external inputs, 2 external ouputs, 2 strict internal
inputs, 2 strict internal outputs (the sets of strict internal inputs and
outputs are the same, but they are labelled differently by $f_{int}$ and $g_{int}$). We
do the same thing with $[\id]$ and notice that the two resulting cospans are
isomorphic. Thus they are equal in the category where cospans are quotiented by
isos.}

We assume that in $[\sym] = [\id]$ it was meant that $[\sym] = [\id \otimes \id]$, otherwise the equality does not typecheck.

By the definition $[\sym]$ is 
\[
    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
\]
represented as
\begin{figure}[h!]
    \[
    \scalebox{0.5}{
        \tikzfig{reviews/sym}
    }
    \]
\end{figure}

And $[\id \otimes \id]$ is \[
    n \xrightarrow{f'_{ext}} n'' \xrightarrow{f'_{int}} \mathcal{G} \xleftarrow{g'_{int}} m'' \xleftarrow{g'_{ext}} m
\]
represented as 
\begin{figure}[h!]
    \[
    \scalebox{0.5}{
        \tikzfig{reviews/id}
    }
    \]
\end{figure}

Clearly, it should be the case that $\{u_1 \mapsto u_1, \ldots u_4 \mapsto u_4\}$ and similarly for $w_i$. I think this does not follow from our definition of isomorphic cospans, so perhaps we should've required that $\alpha$ and $\gamma$ are identities (below).
\[
    \scalebox{0.75}{
        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
    }
\]
Then, the cospans above are not isomorphic. Consider the cases.
\begin{enumerate}
    \item ${v_1 \mapsto v_1}$ and ${v_2 \mapsto v_2}$ under $\beta$. Then $g_{int};\beta(w_4) \not = g'_{int}(w_4)$
    \item ${v_1 \mapsto v_2}$ and ${v_2 \mapsto v_1}$ under $\beta$. Then $f_{int};\beta(w_4) \not = f'_{int}(w_4)$
\end{enumerate}
This means that the corresponding diagram does not commute and therefore the cospans are not isomorphic.

\paragraph{In the proof}
In the first case, it is immediate that $[f] \not = [g]$, since equal terms
have weak normal forms with the same number of components.
\textit{This argument does not seem to be valid in presence of equations.
Consider an equation $a = a + b$ for example, where a and b are constants of the
signature. Then}
\[
    [a] = [a + b] = [a] + [b]
\]

We only consider equations of SMC terms, \textit{i.e.}, with no $+$ on either side of $=$.

\end{document}