### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over the category of semilattices. 
In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.  
The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.

In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string};
% , as sketched in Fig. \ref{fig:app1}(a)
the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX};
% , as sketched in Fig. \ref{fig:app1}(b)
or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes.
% , as sketched in Fig.~\ref{fig:app2}. 
All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure as the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
 % traced comonoid

% \begin{figure}
% \begin{subfigure}{0.45\linewidth}
% \[
% 	\scalebox{0.75}{
% 	\tikzfig{../figures/conclusion/example-e-graphs_1}
% 	}
% \]
% \subcaption{\;} 
% \end{subfigure}
% \hfill
% \begin{subfigure}{0.45\linewidth}
% \[
% 	\scalebox{0.75}{
% 	\tikzfig{../figures/conclusion/example-e-graphs_2}
% 	}
% \]
% \subcaption{\;}
% \end{subfigure}
% \caption{E-(hyper)graphs for functorial boxes and the ZX-calculus}
% \label{fig:app1}
% \end{figure}

% \begin{figure}
% 	\[
% 	\scalebox{0.65}{
% 	\tikzfig{../figures/conclusion/example-traced}
% 	}
% \]
% 	\caption{E-graphs with cycles and e-hypergraphs with trace}
% 	\label{fig:app2}
% \end{figure}

% In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.  Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}. % traced comonoid
%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments. Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.  However, as we have seen in this paper, the technical details can be delicate. 
% Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}. Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.