### 2025-01-23 LICS_2025/e_graphs_with_bindings/sections/soundness.tex

\section{Soundness and completeness}

In this section we explore soundness and completeness of DPOI rewriting introduced in previous section with respect to rewriting of closed $\Sigma^{+}$-terms.
First we recall the rewriting of $\Sigma$-terms modulo SMC laws.

\begin{definition}
    We say that a $\Sigma$-term $f$ rewrites ($\leadsto$) to a $\Sigma$-term $g$ modulo SMC-laws via a rewrite rule $\langle l, r \rangle$ if they are representable as
    \[
    f = c_{1};(id_{k} \otimes l);c_{2} \qquad g = c_{2};(id_{k} \otimes r);c_{2}~.
    \]
    These representations are so-called SMC normal forms.
\end{definition}

This notion generalises to rewriting of closed $\Sigma$-terms by introducing contexts (in string diagrammatic notation).