### 2025-06-12 POPL2026/SlottedNotes.tex

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%

\documentclass[acmsmall,screen, nonacm, anonymous]{acmart}

\newcommand*{\ONECOLUMN}{}

\input{../preamble}
\input{../macros}
\input{../sample.tikzstyles}
\input{../hypergraph.tikzstyles}
\input{../hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
% \setcopyright{acmlicensed}
% \copyrightyear{2024}
% \acmYear{2024}
% \acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.

%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
% \acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\settopmatter{printacmref=false}
\authorsaddresses{}

\begin{document}

\title{Notes on slotted e-graphs}

\author{Anonymous}

%\email{}
%\orcid{}
%\affiliation{
%  \institution{}
%  \streetaddress{}
%  \city{}
%  \state{}
%  \country{
%  \postcode{}
%}

%\renewcommand{\shortauthors}{}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

\maketitle

\section*{Section}

Slotted e-graphs~\cite{slotted-egraphs} are an extension of e-graphs with two main objectives (apart from representing sets of equivalent terms)
\begin{itemize}
  \item Uniquely represent terms that are different up to a `renaming' of free variables, i.e., terms $a + b$ and $c + d$ where $a,b,c,d$ are distinct variables are represented as the same e-node.
  \item Uniquely represent terms that are different up to renaming of bound variables, i.e., terms $\text{bind}\;\$s_1\; t_1$ and $\text{bind}\; \$s_2\;t_2$ are represented by the same e-node if $(s_1 \; s') \text{bind}\; \$s_1\;t_1 = (s_2 \; s') \text{bind}\;\$s_2\;t_2$ where $s'$ is a fresh name.
\end{itemize}

The first is achieved by parameterising each e-class by the ordered set of free variables (\textit{slots}) the terms of the e-class contain.
For example, a term $a + b$ contains two free variables $a$ and $b$ and therefore the e-class that contains this term is parameterised by two variables.
Non duplication of $a + b$ and $c + d$ is achieved by computing normal forms (\textit{shapes}) of such terms.
In this particular example the shape is $\$0 + \$1$.
One of the components of a slotted e-graph is a hash-map that maps such shapes to e-class id's that contain equivalent terms.
Since two terms evaluate to the same shape, they correspond to a singe entry in an e-graph.

The other one is achieved by both computing the shape and generating a fresh slot each time a term with a binder is added to an e-graph.
This makes it impossible to write a capturing $\beta$-reduction such as $(\lambda x . \lambda y . x) y$.

Both of the above objectives can be achieved by using string diagrams for Cartesian closed SMCs by essentially performing a closure conversion of every functional symbol.

Consider the example in Figure~\ref{fig:slotted_a_b_c_d} where subfigure (c) represents a term $(a + b) + (c + d)$ and note how subterms for $(a+b)$ and $(c + d)$ are shared.
The headers of e-classes specify which free variables parameterise the e-class.
Subfigure (d) represents an e-graph after an equivalent term $a + (b + (c + d))$ was added.

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_example}
  \caption{Slotted e-graph example~\cite{slotted-egraphs}}
  \label{fig:slotted_a_b_c_d}
\end{figure}

We can express the same data using string diagrams as in Figure~\ref{fig:slotted_a_b_c_d_string_1}.
Note that there is no 1-to-1 corresponding between the slotted e-graph and the resulting string diagram (unlike in Figure~\ref{fig:slotted_a_b_c_d_string_2} which can not be obtained by applying rewrite rules to the initial string diagram).
However, the e-box in red roughly corresponds to the invocations $(a)\;(b,c,d)$ and $(a,b)\;(c,d)$ of the slotted e-graph.
We can decompose a string diagram in such a way because of the copy-delete rule.
Also note that we can technically achieve more sharing by not only doing closure conversion on subtrees but on subgraphs as well, i.e., we could share the $\lambda x y . x + y$ 3 times instead of 2.

In the presence of $\beta$-reduction rule we need a richer structure for our string diagrams though.
Consider a term $(\lambda\; x\; .\; f\;x)\; y$ which evaluates to $f \; y$ under $\beta$-reduction.
The corresponding slotted e-graph (after a $\beta$-reduction has been applied) can be seen in Figure~\ref{fig:f_x_y_slotted}.
Note that it contains a cycle (the rightmost \texttt{app} e-node references \texttt{bind} and the latter references the e-class containing \texttt{app} again).
This e-graph contains an infinite family of terms: $(\lambda\; x\; .\; f\;x)\; y$, $f \; y$, $(\lambda\; x\; .\; (\lambda z\; .\; f\;z) x\; y$ and so on.
To fully support this, we need a trace.
An example could be seen in Figure~\ref{fig:f_x_y_string_traced}.

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_string_1_2}
  \caption{String diagram example}
  \label{fig:slotted_a_b_c_d_string_1}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_string_2}
  \caption{String diagram example}
  \label{fig:slotted_a_b_c_d_string_2}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/f_x_y_slotted}
  \caption{Slotted e-graph for $\lambda x\; .\; f\;x$}
  \label{fig:f_x_y_slotted}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/f_x_y_string_traced}
  \caption{String diagram example}
  \label{fig:f_x_y_string_traced}
\end{figure}

\bibliographystyle{acm}
\bibliography{../bibliography}

\end{document}