### 2025-07-04 POPL2026/SlottedNotes.tex
notes update

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%

\documentclass[acmsmall,screen, nonacm, anonymous]{acmart}

\newcommand*{\ONECOLUMN}{}

\input{../preamble}
\input{../macros}
\input{../sample.tikzstyles}
\input{../hypergraph.tikzstyles}
\input{../hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

\usepackage{minted}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
% \setcopyright{acmlicensed}
% \copyrightyear{2024}
% \acmYear{2024}
% \acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.

%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
% \acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\settopmatter{printacmref=false}
\authorsaddresses{}

\begin{document}

\title{Notes on slotted e-graphs}

\author{Anonymous}

%\email{}
%\orcid{}
%\affiliation{
%  \institution{}
%  \streetaddress{}
%  \city{}
%  \state{}
%  \country{
%  \postcode{}
%}

%\renewcommand{\shortauthors}{}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

\maketitle

\section*{Section}

Slotted e-graphs~\cite{slotted-egraphs} are an extension of e-graphs with two main objectives (apart from representing sets of equivalent terms)
\begin{itemize}
  \item Uniquely represent terms that are different up to a `renaming' of free variables, i.e., terms $a + b$ and $c + d$ where $a,b,c,d$ are distinct variables are represented as the same e-node.
  \item Uniquely represent terms that are different up to renaming of bound variables, i.e., terms $\text{bind}\;\$s_1\; t_1$ and $\text{bind}\; \$s_2\;t_2$ are represented by the same e-node if $(s_1 \; s') \text{bind}\; \$s_1\;t_1 = (s_2 \; s') \text{bind}\;\$s_2\;t_2$ where $s'$ is a fresh name.
\end{itemize}

The first is achieved by parameterising each e-class by the ordered set of free variables (\textit{slots}) the terms of the e-class contain.
For example, a term $a + b$ contains two free variables $a$ and $b$ and therefore the e-class that contains this term is parameterised by two variables.
Non duplication of $a + b$ and $c + d$ is achieved by computing normal forms (\textit{shapes}) of such terms.
In this particular example the shape is $\$0 + \$1$.
One of the components of a slotted e-graph is a hash-map that maps such shapes to e-class id's that contain equivalent terms.
Since two terms evaluate to the same shape, they correspond to a singe entry in an e-graph.

The other one is achieved by both computing the shape and generating a fresh slot each time a term with a binder is added to an e-graph.
This makes it impossible to write a capturing $\beta$-reduction such as $(\lambda x . \lambda y . x) y$.

Both of the above objectives can be achieved by using string diagrams for Cartesian closed SMCs by essentially performing a closure conversion of every functional symbol.

Consider the example in Figure~\ref{fig:slotted_a_b_c_d} where subfigure (c) represents a term $(a + b) + (c + d)$ and note how subterms for $(a+b)$ and $(c + d)$ are shared.
The headers of e-classes specify which free variables parameterise the e-class.
Subfigure (d) represents an e-graph after an equivalent term $a + (b + (c + d))$ was added.

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_example}
  \caption{Slotted e-graph example~\cite{slotted-egraphs}}
  \label{fig:slotted_a_b_c_d}
\end{figure}

We can express the same data using string diagrams as in Figure~\ref{fig:slotted_a_b_c_d_string_1}.
Note that there is no 1-to-1 corresponding between the slotted e-graph and the resulting string diagram (unlike in Figure~\ref{fig:slotted_a_b_c_d_string_2} which can not be obtained by applying rewrite rules to the initial string diagram).
However, the e-box in red roughly corresponds to the invocations $(a)\;(b,c,d)$ and $(a,b)\;(c,d)$ of the slotted e-graph.
We can decompose a string diagram in such a way because of the copy-delete rule.
Also note that we can technically achieve more sharing by not only doing closure conversion on subtrees but on subgraphs as well, i.e., we could share the $\lambda x y . x + y$ 3 times instead of 2.

In the presence of $\beta$-reduction rule we need a richer structure for our string diagrams though.
Consider a term $(\lambda\; x\; .\; f\;x)\; y$ which evaluates to $f \; y$ under $\beta$-reduction.
The corresponding slotted e-graph (after a $\beta$-reduction has been applied) can be seen in Figure~\ref{fig:f_x_y_slotted}.
Note that it contains a cycle (the rightmost \texttt{app} e-node references \texttt{bind} and the latter references the e-class containing \texttt{app} again).
This e-graph contains an infinite family of terms: $(\lambda\; x\; .\; f\;x)\; y$, $f \; y$, $(\lambda\; x\; .\; (\lambda z\; .\; f\;z) x\; y$ and so on.
To fully support this, we need a trace.
An example could be seen in Figure~\ref{fig:f_x_y_string_traced}.

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_string_1_2}
  \caption{String diagram example}
  \label{fig:slotted_a_b_c_d_string_1}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_string_2}
  \caption{String diagram example}
  \label{fig:slotted_a_b_c_d_string_2}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/f_x_y_slotted}
  \caption{Slotted e-graph for $\lambda x\; .\; f\;x$}
  \label{fig:f_x_y_slotted}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/f_x_y_string_traced}
  \caption{String diagram example}
  \label{fig:f_x_y_string_traced}
\end{figure}

\subsection{E-matching}

E-matching is a function $match(p,m * a, \beta, m_p)$ that takes in a pattern $p$, an applied-id $m*a$ to match against, a subtitution from variables to applied-ids $\beta$ and a renaming $m_p$ that maps slots of an e-graph to slots of the pattern.

\begin{listing}
\begin{minted}[linenos, fontsize=\small, breaklines]{rust}
  struct State {
    // uses egraph slots.
    partial_subst: Subst, // type Subst = HashMap<String, AppliedId>

    // maps from the pattern slots to the e-graph slots.
    partial_slotmap: SlotMap, //type SlotMap = HashMap<Slot,Slot>
  }

  pub fn ematch_all<L: Language, N: Analysis<L>>(
    eg: &EGraph<L, N>,
    pattern: &Pattern<L>,
) -> Vec<Subst> {
    let mut out = Vec::new();
    for i in eg.ids() {
        let i = eg.mk_sem_identity_applied_id(i); // applies identity renaming to an id to turn it into AppliedId
        out.extend(
            ematch_impl(pattern, State::default(), i, eg) // this is `match'
                .into_iter()
                .map(final_subst),
        );
    }
    out
}

\end{minted}
\end{listing}

\begin{listing}
\begin{minted}[linenos, fontsize=\small, breaklines]{rust}
fn ematch_impl<L: Language, N: Analysis<L>>(
    pattern: &Pattern<L>,
    st: State,
    i: AppliedId,
    eg: &EGraph<L, N>,
) -> Vec<State> {
    match &pattern {
        Pattern::PVar(v) => {
            let mut st = st;
            if let Some(j) = st.partial_subst.get(v) {
                if !eg.eq(&i, j) {
                    return Vec::new();
                }
            } else {
                st.partial_subst.insert(v.clone(), i);
            }
            vec![st]
        }
        Pattern::ENode(n, children) => {
            let mut out = Vec::new();
            for nn in eg.enodes_applied(&i) { //traverses e_nodes of the e-class
                let d = std::mem::discriminant(n);
                let dd = std::mem::discriminant(&nn);
                if d != dd {
                    continue;
                };

                //takes the e_node that matches 'n'.
                ematch_node(&st, eg, &n, children, &mut out, &nn);
            }
            out
        }
        Pattern::Subst(..) => panic!(),
    }
}

\end{minted}
\end{listing}

\begin{listing}
\begin{minted}[linenos, fontsize=\small, breaklines]{rust}
fn ematch_node<L: Language, N: Analysis<L>>(
    st: &State,
    eg: &EGraph<L, N>,
    n: &L, // from the pattern
    children: &[Pattern<L>],
    out: &mut Vec<State>,
    nn: &L, //candidate e_node from e-class
) {
    'nodeloop: for n2 in eg.get_group_compatible_weak_variants(&nn) {
        if CHECKS {
            assert_eq!(&nullify_app_ids(n), n);
        }

        let clear_n2 = nullify_app_ids(&n2);
        // We can use weak_shape here, as the inputs are nullified
        // i.e. they only have id0() without slot args, so there are no permutations possible.
        let (n_sh, _) = n.weak_shape();
        let (clear_n2_sh, _) = clear_n2.weak_shape();
        if n_sh != clear_n2_sh {
            continue 'nodeloop;
        } //checks if the shapes are equal, for example lambda x . f and lambda y . f have the same shape

        let mut st = st.clone();

        for (x, y) in clear_n2
            .all_slot_occurrences()
            .into_iter()
            .zip(n.all_slot_occurrences().into_iter())
        {
          // checks if mapping m_p is not ambiguous
            if !try_insert_compatible_slotmap_bij(x, y, &mut st.partial_slotmap) {
                continue 'nodeloop;
            }
        }

        let mut acc = vec![st];
        // recursively matches children
        for (sub_id, sub_pat) in n2.applied_id_occurrences().into_iter().zip(children.iter()) {
            let mut next = Vec::new();
            for a in acc {
                next.extend(ematch_impl(sub_pat, a, sub_id.clone(), eg));
            }
            acc = next;
        }

        out.extend(acc);
    }
}

\end{minted}
\end{listing}

\section{Explicit substitution}

Lambda calculus with explicit substitution (also $\lambda_{\sigma}$-calculus) made the substitution operator a first-class element of the calculus (rather than a meta-level operation) to bridge the gap between lambda calculus and its concrete implementations~\cite{Abadi_Cardelli_Curien_Lévy_1991}.
However, the existence of unique normal forms turned out to be a delicate matter and to achieve uniqueness, lambda terms must be quotiented by \textit{graphical} equivalence in addition to $\alpha$-equivalence~\cite{Standardisation}.
We claim that using e-hypergraphs over e-graphs (or slotted e-graphs) may be more beneficial for doing equational reasoning for lambda calculus with explicit substitution.

The rules can be seen in a table below (plust a standard rule for $(\lambda x . u)v = u[x / v]$)

\begin{tabular}{lcc}
\texttt{subst\_same\_var} & $x[x / u] = u$ &\\
\texttt{subst\_diff\_var} & $y[x / u] = y$ & if $x \not \in \mathcal{F}(u)$\\
\texttt{subst\_application} & $(uu')[x / v] = u[x / v]u'[x / v]$ & \\
\texttt{subst\_lambda} & $(\lambda x . u)[y / v] = \lambda x . (u[y / v])$ & if $x \not \in \mathcal{F}(v)$ \\
\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$\\
\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])[x / v[x / v']]$ & if ???\\
\texttt{subst\_comm} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v]$ & if $x \not \in \mathcal{F}(v')$ and $x' \not \in \mathcal{F}(v)$\\
\end{tabular}

while such rules are absorbed when using string diagrams (or hypergraphs) akin to proof nets, they are not absorbed by e-graph representation.
Below one could see the number of e-nodes for different set of equations enabled for saturating an expression \texttt{plus 2 2} using Church encoding.

\begin{tabular}{lcccc}
  Set of equations & built-in subst & explicit subst & explicit subst & explicit subst\\
  &&& + interchange&  + interchange\\
  &&&&  + comm\\
  \# of e-nodes & 40 & 161 & 289 & 356
\end{tabular}

String diagrams represent the same equivalence without loss of generality and without handling equations for substitutions explicitly.

\bibliographystyle{acm}
\bibliography{../bibliography}

\end{document}